{\small Filamentary transport in high-power H-mode conditions and in no/small-ELM regimes to predict heat and particle loads on PFCs for future devices}

### 2017-11-13 Meetings/GPMNov2017/gpm_november2017_topic21.tex
added further slides to GPM

\begin{frame}{L-Mode analysis: I$_p$ scan at constant q$_{95}$}
\Fontvi
  \vspace{-1cm}
  \begin{columns}    
  \begin{column}{0.65\textwidth}
    \centering{\includegraphics<1>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/EquilibraLparallelConstantQ95}}
    \includegraphics<2>[height=.81\textheight]{../../Experiments/AUG/analysis/pdfbox/GeneralIpScanConstantq95}
    \centering{\includegraphics<3>[height=.85\textheight]{../../Experiments/TCV/analysis/pdfbox/EquilibriaLparallelConstantQ95}}
    \centering{\includegraphics<4>[height=.9\textheight]{../../Experiments/TCV/analysis/pdfbox/CurrentScanConstantQ95}}
    \centering{\includegraphics<5>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/IpConstantQ95_Profiles_UsDiv}}
    \centering{\includegraphics<6>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/n0_CurrentScanConstantQ95}}
    \centering{\includegraphics<7>[width=.6\textheight]{../../Experiments/AUG/analysis/pdfbox/RadiationPeakDensityIpScan_constantQ95}}
    \centering{\includegraphics<8>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/IpConstantq95_samedensity}}
    \centering{\includegraphics<9>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/CompareTargetProfilesConstantQ95}}
    \centering{\includegraphics<10>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/BlobSizeCurrentScanConstantQ95}}
    \centering{\includegraphics<11>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/GPI_blobs_AUG}}
    \centering{\includegraphics<12>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/LambdaSizeIpScanConstantQ95}}

  \end{column}
  \begin{column}{0.35\textwidth}
    \begin{itemize}
      \item<1|only@1> AUG: All the shots were performed in the so-called
        Edge Optmized Configuration (EOC) shape. 
      \item<1|only@1> AUG: We matched correctly the shape and the L$_{\parallel}$
        here shown from outer divertor plate up to X-point 
      \item<2|only@2> AUG: The scan was performed with similar puffing
        rates for intermediate and higher current (0.8-1
        MA) whereas we reduced it at lower current to avoid early
        disruption
      \item<2|only@2> AUG: The total power (Ohmic plus NBI) was kept
        constant throughout the scan
      \item<2|only@2> AUG: We have comparable edge density, divertor neutral
        pressure and divertor temperature
      \item<3|only@3> TCV: We repeat the same excercise at TCV with a
        slight difference in the profile of parallel connection
        length. This required operation at unusual low toroidal field
        (up to 0.8T)
      \item<4|only@4> TCV: no additional heating
        has been used. Nevertheless the difference in power crossing the separatrix
        is small
      \item<4|only@4> TCV: The difference in target pressure similar
        to AUG behavior. 
      \item<5|only@5> AUG: At comparable edge density Upstream profiles are
        different with the tendency to develop shoulder easier at
        lower current. \alert{We have flattening of the upstream
          profiles only when $\Lambda_{div}$ is well above one on all
          the profile}
      \item<6|only@6> AUG: Neutrals estimated using calibrated D$_{\alpha}$
        cameras coupled with values of density and temperature at the
        target suggest a larger neutral density at lower current (even
        with comparable values of edge electron density)
      \item<7|only@7> AUG: At all the current a clear roll-over in
        peak density observed and we can also track the radiation
        front move away from target earlier in density at lower current
      \item<8|only@8> TCV: This tendency is not observed for TCV where
        profiles seem resilient to modification of Bt even though we
        reached pretty high value of $\Lambda_{div}$ all along the profile.
      \item<9|only@9> TCV: This is due to the fact that we can't
        observed during the density ramp any signature of rollover or detachment,
        \alert{whereas upstream profile modification at TCV are only observed
        well after rollover}
      \item<10|only@10> AUG: There is the tendency towards larger blobs
        at lower current at all values of $\Lambda_{div}$
      \item<11|only@11> AUG: This need to be further confirmed
        independently by GPI measurements. The present setup allows
        for space resolution of the order of mm with a 397 kframe/s
        sampling rate. Small amount of He gas flux needed ($\sim
        1.5e^{17} atoms/ms$)
      \item<12|only@12> TCV: this is not confirmed for TCV where blob
        sizes is rather constant over a rather large range of
        $\Lambda_{div}$. At the same time we do not observe target
        density rollover neither upstream profile flattening 
      \end{itemize}
    \end{column}
\end{columns}
\end{frame}

\begin{frame}{L-Mode analysis: I$_p$ scan at constant B$_{t}$}
\Fontvi
  \vspace{-1cm}
  \begin{columns}    
  \begin{column}{0.65\textwidth}
    \centering{\includegraphics<1>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/EquilibraLparallelConstantBt}}
    \centering{\includegraphics<2>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/GeneralIpScanConstantBt}}
    \centering{\includegraphics<3>[height=.85\textheight]{../../Experiments/TCV/analysis/pdfbox/EquilibriaLparallelConstantBt}}
    \centering{\includegraphics<4>[height=.9\textheight]{../../Experiments/TCV/analysis/pdfbox/CurrentScanConstantBt}}
    \centering{\includegraphics<5>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/IpConstantBt_Profiles_UsDiv}}
    \centering{\includegraphics<6>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/n0_CurrentScanConstantBt}}
    \centering{\includegraphics<7>[width=.65\textheight]{../../Experiments/AUG/analysis/pdfbox/RadiationPeakDensityIpScan_constantBT}}
    \centering{\includegraphics<8>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/IpConstantBt_samedensity}}
    \centering{\includegraphics<9>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/CompareTargetProfilesConstantBt}}
    \centering{\includegraphics<10>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/BlobSizeCurrentScanConstantBT}}
    \centering{\includegraphics<11>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/LambdaSizeIpScanConstantBt}}

@@

\begin{frame}{Fast electron generation}
\Fontvi
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering{\includegraphics[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/SXR_ken}}
    \end{column}
    \begin{column}{0.5\textwidth}
\begin{itemize}
      \item Bursts of nonthermal 2nd harmonic electron cyclotron emission and edge soft X-ray emission often occur at start of ELM filament eruption in low collisionality AUG pulses – evidence of filament reconnection
\item Soft X-ray bursts were seen at start of ELM filament eruption in T21-AUG pulses, but not ECE bursts, probably due to higher collisionality in these pulses
\item Peak nonthermal ECE emission appears to originate from top of pedestal
\item Work is ongoing to determine conditions for occurrence of ECE \& SXR bursts, e.g. in terms of pedestal collisionality – there is no clear correlation between ELM size \& SXR flux
\end{itemize}
\end{column}

@@

\begin{frame}{Code requirements}
  \begin{itemize}
    \item 
  \end{itemize}
\end{frame}

>>>>

\begin{frame}{L-Mode analysis: I$_p$ scan at constant q$_{95}$}
\Fontvi
  \vspace{-1cm}
  \begin{columns}    
  \begin{column}{0.65\textwidth}
    \centering{\includegraphics<1>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/EquilibraLparallelConstantQ95}}
    \includegraphics<2>[height=.81\textheight]{../../Experiments/AUG/analysis/pdfbox/GeneralIpScanConstantq95}
    \centering{\includegraphics<3>[height=.85\textheight]{../../Experiments/TCV/analysis/pdfbox/EquilibriaLparallelConstantQ95}}
    \centering{\includegraphics<4>[height=.9\textheight]{../../Experiments/TCV/analysis/pdfbox/CurrentScanConstantQ95}}
    \centering{\includegraphics<5>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/IpConstantQ95_Profiles_UsDiv}}
    \centering{\includegraphics<6>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/n0_CurrentScanConstantQ95}}
    \centering{\includegraphics<7>[width=.6\textheight]{../../Experiments/AUG/analysis/pdfbox/RadiationPeakDensityIpScan_constantQ95Zoom}}
    \centering{\includegraphics<8>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/IpConstantq95_samedensity}}
    \centering{\includegraphics<9>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/CompareTargetProfilesConstantQ95}}
    \centering{\includegraphics<10>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/BlobSizeCurrentScanConstantQ95}}
    \centering{\includegraphics<11>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/GPI_blobs_AUG}}
    \centering{\includegraphics<12>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/LambdaSizeIpScanConstantQ95}}

  \end{column}
  \begin{column}{0.35\textwidth}
    \begin{itemize}
      \item<1|only@1> AUG: All the shots were performed in the so-called
        Edge Optmized Configuration (EOC) shape. 
      \item<1|only@1> AUG: We matched correctly the shape and the L$_{\parallel}$
        here shown from outer divertor plate up to X-point 
      \item<2|only@2> AUG: The scan was performed with similar puffing
        rates for intermediate and higher current (0.8-1
        MA) whereas we reduced it at lower current to avoid early
        disruption
      \item<2|only@2> AUG: The total power (Ohmic plus NBI) was kept
        constant throughout the scan
      \item<2|only@2> AUG: We have comparable edge density, divertor neutral
        pressure and divertor temperature
      \item<3|only@3> TCV: We repeat the same excercise at TCV with a
        slight difference in the profile of parallel connection
        length. This required operation at unusual low toroidal field
        (up to 0.8T)
      \item<4|only@4> TCV: no additional heating
        has been used. Nevertheless the difference in power crossing the separatrix
        is small
      \item<4|only@4> TCV: The difference in target pressure similar
        to AUG behavior. 
      \item<5|only@5> AUG: At comparable edge density Upstream profiles are
        different with the tendency to develop shoulder easier at
        lower current. \alert{We have flattening of the upstream
          profiles only when $\Lambda_{div}$ is well above one on all
          the profile}
      \item<6|only@6> AUG: Neutrals estimated using calibrated D$_{\alpha}$
        cameras coupled with values of density and temperature at the
        target suggest a larger neutral density at lower current (even
        with comparable values of edge electron density)
      \item<7|only@7> AUG: At all the current a clear roll-over in
        peak density observed and we can also track the radiation
        front move away from target earlier in density at lower current
      \item<8|only@8> TCV: This tendency is not observed for TCV where
        profiles seem resilient to modification of Bt even though we
        reached pretty high value of $\Lambda_{div}$ all along the profile.
      \item<9|only@9> TCV: This is due to the fact that we can't
        observed during the density ramp any signature of rollover or detachment,
        \alert{whereas upstream profile modification at TCV are only observed
        well after rollover}
      \item<10|only@10> AUG: There is the tendency towards larger blobs
        at lower current at all values of $\Lambda_{div}$
      \item<11|only@11> AUG: This need to be further confirmed
        independently by GPI measurements (I. Cziegler \textit{et al.}). The present setup allows
        for space resolution of the order of mm with a 397 kframe/s
        sampling rate. Small amount of He gas flux needed ($\sim
        1.5e^{17} atoms/ms$)
      \item<12|only@12> TCV: this is not confirmed for TCV where blob
        sizes is rather constant over a rather large range of
        $\Lambda_{div}$. At the same time we do not observe target
        density rollover neither upstream profile flattening 
      \end{itemize}
    \end{column}
\end{columns}
\end{frame}

\begin{frame}{L-Mode analysis: I$_p$ scan at constant B$_{t}$}
\Fontvi
  \vspace{-1cm}
  \begin{columns}    
  \begin{column}{0.65\textwidth}
    \centering{\includegraphics<1>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/EquilibraLparallelConstantBt}}
    \centering{\includegraphics<2>[height=.8\textheight]{../../Experiments/AUG/analysis/pdfbox/GeneralIpScanConstantBt}}
    \centering{\includegraphics<3>[height=.85\textheight]{../../Experiments/TCV/analysis/pdfbox/EquilibriaLparallelConstantBt}}
    \centering{\includegraphics<4>[height=.9\textheight]{../../Experiments/TCV/analysis/pdfbox/CurrentScanConstantBt}}
    \centering{\includegraphics<5>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/IpConstantBt_Profiles_UsDiv}}
    \centering{\includegraphics<6>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/n0_CurrentScanConstantBt}}
    \centering{\includegraphics<7>[width=.65\textheight]{../../Experiments/AUG/analysis/pdfbox/RadiationPeakDensityIpScan_constantBTZoom}}
    \centering{\includegraphics<8>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/IpConstantBt_samedensity}}
    \centering{\includegraphics<9>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/CompareTargetProfilesConstantBt}}
    \centering{\includegraphics<10>[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/BlobSizeCurrentScanConstantBT}}
    \centering{\includegraphics<11>[width=\textwidth]{../../Experiments/TCV/analysis/pdfbox/LambdaSizeIpScanConstantBt}}

@@

\begin{frame}{Fast electron generation (K. McClements \textit{et al})}
\Fontvi
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering{\includegraphics[width=\textwidth]{../../Experiments/AUG/analysis/pdfbox/SXR_ken}}
    \end{column}
    \begin{column}{0.5\textwidth}
\begin{itemize}
      \item Bursts of nonthermal 2nd harmonic electron cyclotron emission and edge soft X-ray emission often occur at start of ELM filament eruption in low collisionality AUG pulses – evidence of filament reconnection
\item Soft X-ray bursts were seen at start of ELM filament eruption in T21-AUG pulses, but not ECE bursts, probably due to higher collisionality in these pulses
\item Peak nonthermal ECE emission appears to originate from top of pedestal
\item Work is ongoing to determine conditions for occurrence of ECE \& SXR bursts, e.g. in terms of pedestal collisionality – there is no clear correlation between ELM size \& SXR flux
\end{itemize}
\end{column}

@@

\begin{frame}{Code requirements}
  \begin{itemize}
    \item Dedicated simulations with first principles codes (GBS,
      HESEL)
    \item Check if trends can be reproduced
    \item potential input for future experiments regarding role of
        f.x. neutrals and recycling
  \end{itemize}
\end{frame}
